Transferability of clinical prediction models for early trauma care

### 2019-09-07 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

## TABLE SETUP
  # Table 1 = Sample characteristics table
  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
  names(high.low) <- c("Characteristic", "Level", "High volume", "Low volume")
  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
  names(met.nonmet) <- c("Characteristic", "Level", "Metropolitan", "Non-metropolitan")
  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
  names(multi.single) <- c("Characteristic", "Level", "Multiple centres", "Single centre")
   # Create complete table one (test)
   complete.table.one <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$non.metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$multi.centre)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$single.centre.11001)
   names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
   complete.table.one$characteristic <- c("n", "Age", "Gender", "", "GCS", "SBP", "RR", "30DS", "", "ISS", "NISS", "ISS>15", "")
  # Table 2 = result.tables, cropped to only show validation
  table.2 <- results.all.data[1:3,]
  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
   # Create complete table two (test) 
   complete.table.two <- results.all.data[1:3,]
   rownames(complete.table.two) <- c("mistriage", "undertriage", "overtriage")
  # Table 3 = Cropped to only show transfer
  table.3 <- results.all.data[4:6,]
  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 4 = Cropped to only show transfer differance
  table.4 <- results.all.data[7:9,]
  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
```
---
abstract: |

@@

  **Conclusions:** In all samples, model transfer was associated with high 
  levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
  sample to the other sample in the data set led to an increased mistriage
  rate, primarily due to increased overtriage.

@@

### Development and validation sample
The samples in each data set were split into two subsamples using a temporal
split based on the date of trauma. The earlier subsample was the development
sample, and the later subsample was the validation sample. The development
sample contained 75 events (events being patients who died within 30 days of the
trauma) and all non-events (non-events being patients who survived 30 days past
the trauma) during the same time. The rationale for including 75 events was that
25 events per variable have been shown to result in less biased coefficient
estimates compared to the more standard 10 events per variable [@RN30;
@Courvoisier2011]. The validation sample contained 100 events and at least 100
non-events, which was suggested as the minimum number by Vergouwe in 2005 for
external validation samples [@RN37]. Figure 1 shows the development of the
samples using the high- and low-volume centre data set as an example. The
minimum sample sizes of the development and validation samples were 140 and 200,
respectively.  We performed analyses only on data sets for which all samples
included at least the minimum sample size.

@@

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In all data sets, model transfer between samples resulted in
substantial uncertainty in terms of its effect on model performance. When comparing the
transferred performance to the receiving sample validation performance, the
resulting differences also showed considerable uncertainty.

The most notable effects on model performance following model transfer were observed
after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
resulted in an increase in the mistriage rate of 
`r substr(format(highest.comp.mistriage), 1, 5)`.  Mainly contributing to this was
a marked increase in overtriage.  For this transfer, the 95% UI ranged from 
`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty, suggesting
that our findings are compatible with a `r UI.top.bot[1]` decrease to a 
`r UI.top.bot[2]` increase in mistriage. The negative 95% UI of `r UI.top.bot[1]`
would signify an improvement in model performance. The higher 95% UI of 
`r UI.top.bot[2]` would mean a marked increase in model mistriage compared to the
point estimate of `r substr(format(highest.comp.mistriage), 1, 5)`.

Due to centralized trauma care, non-metropolitan centres generally have fewer 
resources and less experience receiving trauma patients. They may rely 
on clinical prediction models more heavily for accurate trauma triage, and 
an increase in overtriage can lead to further strains on already limited 
resources. Increased undertriage could possibly lead to increased patient 
mortality.

Recognizing these risks and updating models accordingly could lead
to increased accuracy during trauma patient triage, potentially saving lives by
minimizing undertriage, and likely saving resources by minimizing overtriage. With
registries such as SweTrau, the analysis of model transfer could be performed
continuously to optimize the models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

@@

## Strengths and limitations
### Strengths
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres and then transferred as clinical
recommendations to minor non-metropolitan centres. The results should therefore
be of practical importance.

@@

## Conclusion
Model transfer resulted in large and varying levels of uncertainty in in terms of its effect
on mistriage. Therefore, model transfer can be unpredictable, potentially leading 
to increased mistriage but in some cases also to a decrease in mistriage. 
Noticeably, transfer of the `r colnames(highest.comp.mistriage)` model led to an 
increased mistriage of `r highest.comp.mistriage`. This observed effect was also 
associated with considerable uncertainty. Recognizing the limitations of this study, 
the authors believe that further studies are warranted due to the potential economic 
costs and patient consequences of poor model transfers.

>>>>

## TABLE SETUP
  # Capitalize names before table creation
  names(results.all.data) <- c("High-volume", "Low-volume", "Metropolitan", "Non-metropolitan", "Multiple centres", "Single centre")
  # Table 1 = Sample characteristics table
  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
  names(high.low) <- c("Characteristic", "Level", "High volume", "Low volume")
  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
  names(met.nonmet) <- c("Characteristic", "Level", "Metropolitan", "Non-metropolitan")
  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
  names(multi.single) <- c("Characteristic", "Level", "Multiple centres", "Single centre")
   # Create complete table one (test)
   complete.table.one <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$non.metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$multi.centre)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$single.centre.11001)
   names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
   complete.table.one$characteristic <- c("n", "Age", "Gender", "", "GCS", "SBP", "RR", "30DS", "", "ISS", "NISS", "ISS>15", "")
  # Table 2 = result.tables, cropped to only show validation
  table.2 <- results.all.data[1:3,]
  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
   # Create complete table two (test) 
   complete.table.two <- results.all.data[1:3,]
   rownames(complete.table.two) <- c("mistriage", "undertriage", "overtriage")
  # Table 3 = Cropped to only show transfer
  table.3 <- results.all.data[4:6,]
  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 4 = Cropped to only show transfer differance
  table.4 <- results.all.data[7:9,]
  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
```
---
abstract: |

@@

  **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
  Models developed in samples with a high number of patients performed the best after transfer. Affirming 
  this, the worst performance after transfer was the low-volume model transferred to the high-volume 
  sample, resulting in a significant mistriage increase of `r highest.comp.mistriage`, this was
  primarily due to increased overtriage.

@@

### Development and validation sample
The samples in each data set were split into two subsamples using a temporal
split based on the date of trauma. The earlier subsample was the development
sample, and the later subsample was the validation sample. The development
sample contained 75 events (events being patients who died within 30 days of the
trauma) and all non-events (non-events being patients who survived 30 days past
the trauma) during the same time. The rationale for including 75 events was that
25 events per variable have been shown to result in less biased coefficient
estimates compared to the more standard 10 events per variable [@RN30;
@Courvoisier2011]. The validation sample contained 100 events and at least 100
non-events, which was suggested as the minimum number by Vergouwe in 2005 for
external validation samples [@RN37]. Figure 1 shows the development of the
samples using the high- and low-volume centre data set as an example. The
minimum sample sizes of the development and validation samples were 150 and 200,
respectively.  We performed analyses only on data sets for which all samples
included at least the minimum sample size.

@@

# Discussion
This study aimed to assess how transfers of clinical prediction models for early
trauma care between different contexts within a single health system affect
mistriage rates. In many data sets, varying degrees of uncertainty were observed 
when comparing the transferred performance to the receiving sample validation performance.

The most notable effect on model performance following model transfer were observed
after transferring the `r colnames(highest.comp.mistriage)` model. This transfer
resulted in an increase in the mistriage rate of 
`r substr(format(highest.comp.mistriage), 1, 5)`. Mainly contributing to this was
an increase in overtriage.  For this transfer, the 95% UI ranged from 
`r UI.top.bot[1]` to `r UI.top.bot[2]`, indicating a high level of uncertainty. 
In contrast, the transfer of the high-volume model to the low-volume sample 
reduced mistriage, this was primarily due to reduced overtriage.

Low-volume centres do by definition receive fewer trauma patients, and therefore 
less data is available to establish accurate clinical prediction models. It is 
reasonable to assume this is the reason for the marked increase in mistriage 
when transferring the low-volume model to the high-volume sample. This also 
explains why the low-volume model and the single-centre model showed the highest 
levels of mistriage during model validation (i.e. when applied to the sample in 
which the model was created). Further confirming this, the transfer of the 
high-volume model to the low-volume sample reduces mistriage.

In a clinical context, this seems to suggest clinical prediction models for early 
trauma care should be developed in settings with a high number of trauma patients. 
These models could then be transferred to settings with less trauma patients. Low-volume 
centres generally have fewer resources and less experience receiving trauma patients, 
these centres could therefore benefit by using models developed in centres with more 
experience receiving trauma patients

Recognizing settings with high mistriage and updating models accordingly could lead
to increased accuracy during trauma patient triage, potentially saving lives by
minimizing undertriage, and likely saving resources by minimizing overtriage. With
registries such as SweTrau, the analysis of model transfer could be performed
continuously to optimize the models currently in use and advise which model
transfers might be detrimental (or potentially beneficial).

@@

## Strengths and limitations
### Strengths
The main strength of this study is that its design realistically reflects
potential model transfers: clinical prediction models for trauma are usually
developed in large metropolitan centres with a high volume of patients and 
then transferred as clinical recommendations to minor low-volume centres. The 
results should therefore be of practical importance.

@@

## Conclusion
Model transfer resulted in varying levels of uncertainty in in terms of its effect
on mistriage. In general, samples with a high number of patients had the lowest mistriage 
both during validation and after transfer to the other sample in the data set. The opposite 
was also true, and the transfer of the `r colnames(highest.comp.mistriage)` 
model led to an increased mistriage of `r highest.comp.mistriage`, noting that this observation 
was associated with considerable uncertainty. Recognizing the limitations of this study, 
the authors believe that further studies are warranted due to the potential economic 
costs and patient consequences of poor model transfers.
